### 2022-03-23 paper/localpsf.tex
wrote lumped mass section

In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$. This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them. This idea was inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$. If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion. Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}). Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.

@@

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise. Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denote the sparse finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
For $u_h \in V_h \subset L^2(\Omega)$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}
Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries $\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i)$ for $i=1,\dots,N$. Here $\mathbb{R}^N_\mathbf{M}$ is the vector space $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$. Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\mu_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\mu}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\mu_h(u) = \boldsymbol{\mu}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\mu_h, \nu_h \in V_h'$. The Riesz representation $\mu_h^* \in V_h$ of $\mu \in V_h'$ becomes $\boldsymbol{\mu}^* = \mathbf{M}^{-1} \boldsymbol{\mu}$. After Galerkin approximation, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ becomes an operator $\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has matrix entries $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$. Notice that $\mathbf{A}$ is dense, and computing one matrix entry $\mathbf{A}_{ij}$ requires an action of $\Aop$, which is already computationally expensive (though we get the entire $j$th column as a side effect of computing $\mathbf{A}_{ij}$). Hence methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$. Computing and storing a matrix representation of $\mathbf{A}$ is out of the question.

If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$, $i=1,\dots,N$. The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes $\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$.

@@

@@

In this section we present three main concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use these moments extensively. Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

\subsection{Impulse response moments}
\label{sec:intromoments}

@@

\subsection{Impulse response support ellipsoids}
\label{eq:mean_and_covariance_estimation}

@@

Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$. This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}). The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.

@@

>>>>

In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$. This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them. This idea was inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$. If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion. Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}). Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.

@@

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise. Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denote the sparse finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
For $u_h \in V_h \subset L^2(\Omega)$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}
Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries $\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i)$ for $i=1,\dots,N$. Here $\mathbb{R}^N_\mathbf{M}$ is the vector space $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$. Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$. The Riesz representation $\rho_h^* \in V_h$ of $\rho \in V_h'$ becomes $\boldsymbol{\rho}^* = \mathbf{M}^{-1} \boldsymbol{\rho}$.

After Galerkin approximation, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ becomes an operator $\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has matrix entries $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$. Notice that $\mathbf{A}$ is dense, and computing one matrix entry $\mathbf{A}_{ij}$ requires an action of $\Aop$, which is already computationally expensive (though we get the entire $j$th column as a side effect of computing $\mathbf{A}_{ij}$). Hence methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$. Computing and storing a matrix representation of $\mathbf{A}$ is out of the question.

If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$ for $i=1,\dots,N$. The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
\begin{equation}
\label{eq:discrete_kernel}
\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
\end{equation}

@@

At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative. However, per \eqref{eq:discrete_kernel}, when we compute impulse responses we are effectively operating with a discrete integral kernel which has $(y,x)$ entry
\begin{equation*}
\boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
\end{equation*}
Even if the discretization $\Aop \rightarrow \mathbf{A}$ preserves integral kernel positivity, the inverse mass matrices $\mathbf{M}^{-1}$ will typically contain negative numbers. If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust. We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation, and using the resulting lumped mass matrix everywhere throughout the approximation method. For typical finite element meshes, lumped mass matrix approximations are typically spectrally equivalent to the mass matrix \cite{MASSLUMP2}. In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other more advanced mass lumping techniques may be used \cite{MASSLUMP2}.

@@

In this section we present two key innovations that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

\subsection{Impulse response moments and support estimation}
\label{sec:intromoments}

@@

%\subsection{Impulse response support ellipsoids}
%\label{eq:mean_and_covariance_estimation}

@@

Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$. This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}). The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.

@@

WORK IN PROGRESS BELOW HERE!!!
